---
title: "Letter Recognition EDA"
author: "Matt Oehler"
date: "March 12, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,comment='')
```

## 1 Description
Being able to autonomously read handwritten documents would be instrumental in saving time for several businesses, and revolutionizing the process of things such as family history. However, since this is a non-trivial task due to the wide variety of handwriting, statisical methods can be used to train a computer to read documents instead of having to do it manually. 

## 2 Data
For this problem we have 16 attirbutes of letters contained in a historical document. The figures below show the structure of the data set, the frequency of each letter occuring in the dataset, and the side by side box plots for each of the covariates for each letter. Some variables vary quite a bit between the letters, but some variables such as 'high' have little variance between the letters

```{r,warning=FALSE,fig.height=3}
setwd("~/Matt BYU/10 Winter 2018/Stat 536/EDA 7 Letters")
letters <- read.csv('letters.txt',header=T)
# library(xtable)
# xtable(head(letters[,1:8]))
# xtable(head(letters[,9:16]))

```
\begin{table}[ht]
\centering
\begin{tabular}{rlrrrrrrr}
  \hline
 & letter & xbox & ybox & width & high & pix & xbar & ybar \\ 
  \hline
  1 & I &   5 &  12 &   3 &   7 &   2 &  10 &   5 \\ 
  2 & D &   4 &  11 &   6 &   8 &   6 &  10 &   6 \\ 
  3 & N &   7 &  11 &   6 &   6 &   3 &   5 &   9 \\ 
  4 & G &   2 &   1 &   3 &   1 &   1 &   8 &   6 \\ 
  5 & S &   4 &  11 &   5 &   8 &   3 &   8 &   8 \\ 
  6 & B &   4 &   2 &   5 &   4 &   4 &   8 &   7 \\ 
   \hline
\end{tabular}
\end{table}

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrrr}
  \hline
 & x2bar & y2bar & xybar & x2ybar & xy2bar & xege & xegevy & yege \\ 
  \hline
  1 &   5 &   4 &  13 &   3 &   9 &   2 &   8 &   4 \\ 
  2 &   2 &   6 &  10 &   3 &   7 &   3 &   7 &   3 \\ 
  3 &   4 &   6 &   4 &   4 &  10 &   6 &  10 &   2 \\ 
  4 &   6 &   6 &   6 &   5 &   9 &   1 &   7 &   5 \\ 
  5 &   6 &   9 &   5 &   6 &   6 &   0 &   8 &   9 \\ 
  6 &   6 &   6 &   7 &   6 &   6 &   2 &   8 &   7 \\ 
   \hline
\end{tabular}
\end{table}

```{r,fig.height=3.5,fig.width=3.5}
# table(letters$letter)
#tabulations with or without severity
# for(i in 2:5){
  # print(i)
table(letters$letter)

for(i in 2:ncol(letters)){
  nam <- names(letters)[i]
  plot(letters$letter,letters[,i],main=nam)
}
  
  # print(table(crash[,i]))#without severity
  # cat('\n')
# }
```


## 3 Method
I think that a method such as support vector machines or neural networks would be appropriate for this problem since they can handle classification problems when the response has more than two classes. 


## 4 Something I don't know
I don't know of other methods to approach multi-level classification problems. I am also unsure what would happen if we had data where there weren't a balanced level of classes. 

